### 2024-03-29 misc/expo/entropy.tex
Done entropy notes

\proof For the first inequality, let $X'$ be a copy of $X$ that is independent of $(Y,Z)$, so that
$$\eqalign{
\dd\{X, Y-Z\} - \dd\{X, Y\}
&= \Eta\{X'-Y+Z\} - {1\over 2}\Eta\{X'\} - {1\over 2} \Eta\{Y-Z\} \cr
&\hskip60pt-\Eta\{X'-Y\} + {1\over 2}\Eta\{X'\} + {1\over 2}\Eta\{Y\}. \cr
&= \Eta\{X'-Y+Z\} - {1\over 2} \Eta\{Y-Z\} \cr
&\hskip110pt-\Eta\{X'-Y\} + {1\over 2}\Eta\{Y\}.\cr
}$$
Lemma~\ref{\lemaone} with $X'$ in place of $X$ and $-Y$ in place of $Y$ yields
$$\Eta\{X'-Y+Z\} - \Eta\{X'-Y\} \le \Eta\{Y-Z\} -\Eta\{Y\},$$
so
$$\dd\{X, Y-Z\} - \dd\{X, Y\} \le {1\over 2} \Eta\{Y-Z\} - {1\over 2}\Eta\{Y\},$$
as desired. The alternate version of the statement follows directly from the definition of Ruzsa distance,
since $Y$ and $Z$ are independent.

For the other inequality, we apply Lemma~\ref{\lemfiveone} with $W = Y-Z$ (and the variable $Z$ in that lemma
set to anything that is independent of $X$) to obtain
$$\eqalign{
\dd\{X; Y\given Y-Z\} - \dd\{X,Y\} &\le {1\over 2} \II\{Y:Y-Z\} \cr
&= {1\over 2}\bigl(\Eta\{Y\} + \Eta\{Y-Z\} - \Eta\{Y, Y-Z\}\bigr) \cr
&= {1\over 2}\bigl(\Eta\{Y\} + \Eta\{Y-Z\} - \Eta\{Y, Z\}\bigr) \cr
&= {1\over 2}\bigl(\Eta\{Y-Z\} - \Eta\{Z\}\bigr),\cr
}$$
where in the last line we used independence of $Y$ and $Z$. Independence also gives the alternative version
of the right-hand-side expression.\slug

@@

\proof The second inequality of Lemma~\ref{\lemfivetwo} (setting $Y \gets Y-Z$ and $Z\gets W$) yields
$$\dd\{X; Y-Z\given Y-Z-W\} - \dd\{X, Y-Z\} \le {\Eta\{Y-Z-W\} - \Eta\{W\}\over 2},$$
and the first inequality (without any change of variables) is
$$\dd\{X, Y-Z\} - \dd\{X, Y\} \le {\Eta\{Y-Z\} - \Eta\{Y\}\over 2}.$$
Adding these two inequalites proves the lemma.\slug

@@

\edef\lematwo{\the\sectcount.\the\thmcount}
\parenproclaim Lemma {\advthm} ({\rm\ref{pfr2023},} Lemma A.2).
Let $X$ and $Y$ be discrete random variables taking values in the same abelian
group. Let $(X_1, Y_1)$ and $(X_2,Y_2)$ be conditionally
independent trials of $(X,Y)$ relative to $X+Y$. Then we have
$$\max\bigl(\Eta\{X_1- X_2\}, \Eta\{X_1-Y_2\}\bigr) \le \Eta\{X+Y\} - \II\{X:Y\}.$$
The right-hand side of this expression can also be written $2\Eta\{X\} + 2\Eta\{Y\} - \ee\{X,Y\}$.

@@

@@

@@

\edef\lemruzsacovering{\the\sectcount.\the\thmcount}
\parenproclaim Lemma {\advthm} (Ruzsa covering lemma). Let $G$ be an abelian group and let $A$ and
$B$ be finite subsets of $Z$. Then there is a set $L\subseteq A$ of size at most $|A+B|/|B|$ such that
$A\subseteq L+B-B$.

@@

\medskip\noindent{\it Proof of Theorem~{\thmpfr} assuming Theorem~{\thmsubgroup}.}\enspace
Let $A\subseteq \FF_2$ and $K\ge 1$
such that $|A+A|\le K|A|$. Let $U_A$ be the uniform distribution on $A$, so that $\Eta\{U_A\} = \lg|A|$.
Letting $U_A'$ and $U_A''$ be two independent copies of $U_A$, the sum $U_A' - U_A''$ belongs to $|A+A|$,
so $\Eta\{U_A' + U_A''\} \le \lg|A+A|$, and we have
$$\dd\{U_A, U_A\} = \Eta\{U_A' + U_A''\} - \Eta\{U_A\} \le \lg|A+A|-\lg|A| \le \lg K,$$
since $U_A' + U_A'' = U_A' - U_A''$ in $\FF_2^n$. By Theorem~{\thmsubgroup}, we can find
$H\subseteq G$ such that
$$2\dd\{ U_A, U_H\} \le C'\lg K,$$
where $U_H$ is taken to be uniform on $H$ independently of $U_A$. Independence and the definition
of Ruzsa distance let us conclude
$$\Eta\{U_A-U_H\} \le {\lg|A| + \lg|H| + C'\lg K\over 2},$$
from which applying Proposition~{\propmaxprob} allows us to procure some $x_0 \in \FF_2^n$ such that
$$\pr\{ U_A - U_H = x_0\} \ge {1\over |A|^{1/2} |H|^{1/2} K^{C'/2}}.$$
Since $U_A$ and $U_H$ are uniformly and independently random on $A$ and $H$ respectively, we rewrite this as
$${\bigl| A\cap (H+x_0)\bigr| \over |A|\cdot|H|}  \ge {1\over |A|^{1/2} |H|^{1/2} K^{C'/2}},$$
whence
$$\bigl| A\cap (H+x_0)\bigr|  \ge {|A|^{1/2} |H|^{1/2} \over K^{C'/2}}.$$
By the Ruzsa covering lemma applied to the sets $A$ and $A\cap (H+x_0)$, there is a set $L\subseteq A$ of size
$$\eqalign{
|L| &\le {\bigl| A + (A\cap (H+x_0))\bigr|\over \bigl| A\cap (H+x_0)\bigr|} \cr
&\le {|A + A|\over \bigl| A\cap (H+x_0)\bigr|} \cr
&\le {K|A|\cdot K^{C'/2} \over |A|^{1/2} |H|^{1/2} } \cr
&= {K^{C'/2+1}|A|^{1/2} \over |H|^{1/2} } \cr
}$$
such that
$$A \subseteq L+\bigl(A\cap (H+x_0)\bigr) - \bigl(A\cap (H+x_0)\bigr) \subseteq L+H.$$
Proposition~{\propabsdist} tells us that
$$\bigl| \lg|H| - \lg|A|\bigr| = \bigl| \Eta\{U_H\} - \Eta\{H_A\}\bigr| \le 2\dd\{ U_A, U_H\} \le C'\lg K,$$
so
$$\max\biggl( {|H|\over |A|}, {|A|\over |H|}\biggr) \le K^{C'}$$
and we have
$$|L| \le K^{C'+1}.$$
If $|H|\le |A|$ then we are done (and the factor of $2$ is unnecessary in the theorem statement),
since $A$ is a subset of $|L|$ cosets of $H$.

If not, then we can pick a
subgroup $H'$ of $H$ with $|A|/2 \le |H'|\le |A|$ by iteratively taking hyperplanes until the condition is met.
The group $H$ may be covered by $|H|/|H'| \le 2|H|/|A|$ translates of $H'$, so $A$ can be covered by
$${2|H|\over |A|}|L| \le {2|H|\over |A|}\cdot {K^{C'/2+1}|A|^{1/2} \over |H|^{1/2} } 
= 2 {C'/2+1} {|H|^{1/2} \over |A|^{1/2}} \le 2K{C'+1}$$
translates of $H'$.\slug

@@

\proof If $X$ is the uniform distribution on a coset $a+H$ of a finite subgroup $H$, then letting $X'$ be an
independent copy of $X$, the random variable $X+X'$ is uniform on the coset $a+a+H$, which has the
same size as $|a+H|$, so $\Eta\{X+X'\} = \Eta\{X\}$, meaning that
$$\dd\{X, -X\} = \Eta\{X+X'\} - {\Eta\{X\}\over 2} - {\Eta\{X'\} \over 2} = 0.$$
On the other hand, suppose that $\Eta\{X+X'\} = \Eta\{X\}$, where $X'$ is an independent copy of $X$.
Then
$$\eqalign{
\Eta\{X+X'\} &= 2\Eta\{X\} - \Eta\{X\} \cr
&= \Eta\{X, X'\} - \Eta\{X'\} \cr
&= \Eta\{X+X', X'\} - \Eta\{X'\} \cr
&= \Eta\{X+X'\given X'\},\cr
}$$
so $X+X'$ is independent of $X'$.

@@

\proof The triangle inequality gives $\dd\{X_1,X_1\} = 0$, and since $-x = x$ in $\FF_2^n$,
we also have $\dd\{X_1, - X_1\} = 0$. By the previous proposition, there is a subgroup $H$ such that
$X_1$ has the same distribution as $H$. So $\dd\{X_1, U_H\} = 0$ and by the triangle inequality once again,
$\dd\{X_2, U_H\} = 0$ as well.\slug

@@

If we can prove this proposition, then we have Theorem~{\thmsubgroup} with a certain constant $C'$ which
we shall make explicit during the proof.

\medskip\noindent{\it Proof of Theorem~{\thmsubgroup} assuming Proposition~{\propmaintau}}.\enspace
We proved earlier that $\P(\FF_2^n)$ is a compact topological space, thus so is its Cartesian product with
itself. By Proposition~{\propruzsacontinuous}, the Ruzsa distance functional is continuous on this space,
which implies that $\tau$ is as well. Hence $\tau$ attains its infimum on the space $\P(\FF_2^n)^2$;
let $X_1$ and $X_2$ be two distributions such that $\tau\{X_1, X_2\}$ is minimal. By the contrapositive
of Proposition~{\propmaintau}, we must have $\dd\{X_1, X_2\} = 0$, so by Lemma~{\lemzerosubgroup},
there exists a subgroup $H$ of $\FF_2^n$ such that
$$\dd\{X_1, U_H\} = \dd\{X_2, U_H\} = 0.$$
This means that
$$\dd\{X_1^*, U_H\} \le \dd\{X_1^*, X_1\} + \dd\{ X_1, U_H\} = \dd\{X_1^*, X_1\}$$
and
$$\dd\{X_1^*, X_1\} \le \dd\{X_1^*, U_H\} + \dd\{U_H, X_1\} = \dd\{X_1^*, U_H\},$$
so $\dd\{X_1^*, U_H\} = \dd\{X_1^*, X_1\}$. The same holds for $X_2^*$ and $X_2$.
Then since $\dd\{X_1, X_2\} = 0$ and by minimality of $\tau\{X_1, X_2\}$, we obtain
$$\eqalign{
\eta\dd\{X_1^*, U_H\} + \eta\dd\{X_2^*, U_H\} &= \eta\dd\{X_1, U_H\} + \eta\dd\{X_2, U_H\} \cr
&= \tau\{X_1, X_2\} \cr
&\le \tau\{X_1^*, X_2^*\} \cr
&= \dd\{X_2^*, X_1^*\} + \eta\dd\{X_1^*, X_2^*\} + \eta\dd\{X_2^*, X_1^*\} \cr
&= (1+2\eta) \dd\{X_1^*, X_2^*\}. \cr
}$$
From our choice of $\eta = 1/9$ we can multiply both sides by $9$ to get
$$\dd\{X_1^*, U_H\} + \dd\{X_2^*, U_H\} \le 11\dd\{X_1^*, X_2^*\},$$
and furthermore, since
$$\bigl| \dd\{X_1, U_H\} - \dd\{X_2, U_H\} \bigr| \le \dd\{X_1, X_2\}$$
by two invocations of the triangle inequality, neither $\dd\{X_1^*, U_H\}$ nor
$\dd\{X_2^*, U_H\}$ can be greater than $6\dd\{X_1^*, X_2^*\}$.\slug

@@

>>>>

\proof For the first inequality, let $X'$ be a copy of $X$ that is independent of $(Y,Z)$, so that
$$\eqalign{
\dd\{X, Y-Z\} - \dd\{X, Y\}
&= \Eta\{X'-Y+Z\} - {1\over 2}\Eta\{X'\} - {1\over 2} \Eta\{Y-Z\} \cr
&\hskip60pt-\Eta\{X'-Y\} + {1\over 2}\Eta\{X'\} + {1\over 2}\Eta\{Y\}. \cr
&= \Eta\{X'-Y+Z\} - {1\over 2} \Eta\{Y-Z\} \cr
&\hskip110pt-\Eta\{X'-Y\} + {1\over 2}\Eta\{Y\}.\cr
}$$
Lemma~{\lemaone} with $X'$ in place of $X$ and $-Y$ in place of $Y$ yields
$$\Eta\{X'-Y+Z\} - \Eta\{X'-Y\} \le \Eta\{Y-Z\} -\Eta\{Y\},$$
so
$$\dd\{X, Y-Z\} - \dd\{X, Y\} \le {1\over 2} \Eta\{Y-Z\} - {1\over 2}\Eta\{Y\},$$
as desired. The alternate version of the statement follows directly from the definition of Ruzsa distance,
since $Y$ and $Z$ are independent.

For the other inequality, we apply Lemma~{\lemfiveone} with $W = Y-Z$ (and the variable $Z$ in that lemma
set to anything that is independent of $X$) to obtain
$$\eqalign{
\dd\{X; Y\given Y-Z\} - \dd\{X,Y\} &\le {1\over 2} \II\{Y:Y-Z\} \cr
&= {1\over 2}\bigl(\Eta\{Y\} + \Eta\{Y-Z\} - \Eta\{Y, Y-Z\}\bigr) \cr
&= {1\over 2}\bigl(\Eta\{Y\} + \Eta\{Y-Z\} - \Eta\{Y, Z\}\bigr) \cr
&= {1\over 2}\bigl(\Eta\{Y-Z\} - \Eta\{Z\}\bigr),\cr
}$$
where in the last line we used independence of $Y$ and $Z$. Independence also gives the alternative version
of the right-hand-side expression.\slug

@@

\proof The second inequality of Lemma~{\lemfivetwo} (setting $Y \gets Y-Z$ and $Z\gets W$) yields
$$\dd\{X; Y-Z\given Y-Z-W\} - \dd\{X, Y-Z\} \le {\Eta\{Y-Z-W\} - \Eta\{W\}\over 2},$$
and the first inequality (without any change of variables) is
$$\dd\{X, Y-Z\} - \dd\{X, Y\} \le {\Eta\{Y-Z\} - \Eta\{Y\}\over 2}.$$
Adding these two inequalities proves the lemma.\slug

@@

\edef\lematwo{\the\sectcount.\the\thmcount}
\parenproclaim Lemma {\advthm} ({\rm\ref{pfr2023},} Lemma A.2).
Let $X$ and $Y$ be discrete random variables taking values in the same abelian
group. Let $(X_1, Y_1)$ and $(X_2,Y_2)$ be conditionally
independent trials of $(X,Y)$ relative to $X+Y$. Then we have
$$\max\bigl(\Eta\{X_1- X_2\}, \Eta\{X_1-Y_2\}\bigr) \le \Eta\{X+Y\} + 2\II\{X:Y\}.$$
The right-hand side of this expression can also be written $2\Eta\{X\} + 2\Eta\{Y\} - \ee\{X,Y\}$.

@@

Let us also extract the specific consequence of Lemma~{\lematwo} that we need for the polynomial
Freiman--Ruzsa theorem.

\edef\lematwopfr{\the\sectcount.\the\thmcount}
\proclaim Lemma \advthm. Let $X$ and $Y$ be $G$-valued random variables, let $Z = X+Y$, and let $C$
denote the support of $Z$. Then
$$\eqalign{
\sum_{z\in C} \pr\{Z=z\} \dd\bigl\{ &(X\given Z=z); (Y\given Z=z)\bigr\} \cr
&\le 2\II\{X:Y\}+2\Eta\{Z\}-\Eta\{X,Y\}. \cr
}$$

\proof Let $(X_1, Y_1)$ and $(X_2, Y_2)$ be as in the proof of Lemma~{\lematwo}. Note that
$X_1$ and $Y_2$ are conditionally independent copies of $X$ and $Y$ relative to $Z$, so it suffices
to show that
\edef\eqxoneytwoa{\the\eqcount}
$$\eqalign{
\Eta\{X_1 - Y_2\given Z\} - {1\over 2} &\Eta\{X_1\given Z\} - {1\over 2} \Eta\{Y_2\given Z\} \cr
&\le 2\II\{X:Y\}+2\Eta\{Z\}-\Eta\{X,Y\}. \cr
}\adveq$$
Lemma~{\lematwo} tells us that $\Eta\{X_1 - Y_2\} \le \Eta\{Z\} + 2\II\{X:Y\}$, so that
\edef\eqxoneytwob{\the\eqcount}
$$\Eta\{X_1 - Y_2\given Z\} \le \Eta\{Z\} + 2\II\{X:Y\},\adveq$$
since conditioning does not increase entropy. Then
we expand
\edef\eqxoneytwoc{\the\eqcount}
$$\Eta\{X_1\given Z\} = \Eta\{X_1, X_1 + Y_1\} - \Eta\{Z\} = \Eta\{X,Y\} - \Eta\{Z\},\adveq$$
and $\Eta\{Y_2\given Z\}$ is equal to this quantity as well. Subtracting~\refeq{\eqxoneytwoc}
from~\refeq{\eqxoneytwob} gives us~\refeq{\eqxoneytwoa}, which completes the proof.\slug

@@

The reason we need the factor of two is that if $A$ is almost all of a subgroup, then $K$ is very close
to $1$ and the largest
subgroup of $\FF_2^n$ has cardinality just slightly above half that of $A$. That creates the possibility
that $K^C$ is much less than $2$, making it impossible to satisfy the statement of the conjecture.
Adding the factor of two solves this issue because if $A\subseteq H_0$ with $|A| \ge (1-\eps)|H_0|$, then
we may take a subgroup $H$ of $H_0$ of index $2$ and cover $A$ by two cosets of $H$.

@@

\edef\lemruzsacovering{\the\sectcount.\the\thmcount}
\parenproclaim Lemma {\advthm} (Ruzsa covering lemma). Let $G$ be an abelian group and let $A$ and
$B$ be finite subsets of $G$. Then there is a set $L\subseteq A$ of size at most $|A+B|/|B|$ such that
$A\subseteq L+B-B$.

@@

\medskip\noindent{\it Proof of Theorem~{\thmpfr} assuming Theorem~{\thmsubgroup}.}\enspace
Let $A\subseteq \FF_2^n$ and $K\ge 1$
such that $|A+A|\le K|A|$. Let $U_A$ be the uniform distribution on $A$, so that $\Eta\{U_A\} = \lg|A|$.
Letting $U_A'$ and $U_A''$ be two independent copies of $U_A$, the sum $U_A' - U_A''$ belongs to $|A+A|$,
so $\Eta\{U_A' + U_A''\} \le \lg|A+A|$, and we have
$$\dd\{U_A, U_A\} = \Eta\{U_A' + U_A''\} - \Eta\{U_A\} \le \lg|A+A|-\lg|A| \le \lg K,$$
since $U_A' + U_A'' = U_A' - U_A''$ in $\FF_2^n$. By Theorem~{\thmsubgroup}, we can find
$H\subseteq G$ such that
$$2\dd\{ U_A, U_H\} \le C'\lg K,$$
where $U_H$ is taken to be uniform on $H$ independently of $U_A$. Independence and the definition
of Ruzsa distance let us conclude
$$\Eta\{U_A-U_H\} \le {\lg|A| + \lg|H| + C'\lg K\over 2},$$
from which applying Proposition~{\propmaxprob} allows us to procure some $x_0 \in \FF_2^n$ such that
$$\pr\{ U_A - U_H = x_0\} \ge {1\over |A|^{1/2} |H|^{1/2} K^{C'/2}}.$$
Since $U_A$ and $U_H$ are uniformly and independently random on $A$ and $H$ respectively, we rewrite this as
$${\bigl| A\cap (H+x_0)\bigr| \over |A|\cdot|H|}  \ge {1\over |A|^{1/2} |H|^{1/2} K^{C'/2}},$$
whence
$$\bigl| A\cap (H+x_0)\bigr|  \ge {|A|^{1/2} |H|^{1/2} \over K^{C'/2}}.$$
By the Ruzsa covering lemma applied to the sets $A$ and $A\cap (H+x_0)$, there is a set $L\subseteq A$ of size
$$\eqalign{
|L| &\le {\bigl| A + (A\cap (H+x_0))\bigr|\over \bigl| A\cap (H+x_0)\bigr|} \cr
&\le {|A + A|\over \bigl| A\cap (H+x_0)\bigr|} \cr
&\le {K|A|\cdot K^{C'/2} \over |A|^{1/2} |H|^{1/2} } \cr
&= {K^{C'/2+1}|A|^{1/2} \over |H|^{1/2} } \cr
}$$
such that
$$A \subseteq L+\bigl(A\cap (H+x_0)\bigr) - \bigl(A\cap (H+x_0)\bigr) \subseteq L+H.$$
Proposition~{\propabsdist} tells us that
$$\bigl| \lg|H| - \lg|A|\bigr| = \bigl| \Eta\{U_H\} - \Eta\{H_A\}\bigr| \le 2\dd\{ U_A, U_H\} \le C'\lg K,$$
so
$$\max\biggl( {|H|\over |A|}, {|A|\over |H|}\biggr) \le K^{C'}$$
and we have
$$|L| \le K^{C'+1}.$$
If $|H|\le |A|$ then we are done (and the factor of $2$ is unnecessary in the theorem statement),
since $A$ is a subset of $|L|$ cosets of $H$.

If not, then we can pick a
subgroup $H'$ of $H$ with $|A|/2 \le |H'|\le |A|$ by iteratively taking hyperplanes until the condition is met.
The group $H$ may be covered by $|H|/|H'| \le 2|H|/|A|$ translates of $H'$, so $A$ can be covered by
$${2|H|\over |A|}|L| \le {2|H|\over |A|}\cdot {K^{C'/2+1}|A|^{1/2} \over |H|^{1/2} } 
= 2K^{C'/2+1} {|H|^{1/2} \over |A|^{1/2}} \le 2K^{C'+1}$$
translates of $H'$.\slug

@@

\proof If $X$ is the uniform distribution on a coset $a+H$ of a finite subgroup $H$, then letting $X'$ be an
independent copy of $X$, the random variable $X+X'$ is uniform on the coset $a+a+H$, which has the
same size as $|a+H|$, so $\Eta\{X+X'\} = \Eta\{X\}$, meaning that
$$\dd\{X, -X\} = \Eta\{X+X'\} - {\Eta\{X\}\over 2} - {\Eta\{X'\} \over 2} = 0.$$
On the other hand, suppose that $\Eta\{X+X'\} = 2\Eta\{X\}$, where $X'$ is an independent copy of $X$.
Then
$$\eqalign{
\Eta\{X+X'\} &= 2\Eta\{X\} - \Eta\{X\} \cr
&= \Eta\{X, X'\} - \Eta\{X'\} \cr
&= \Eta\{X+X', X'\} - \Eta\{X'\} \cr
&= \Eta\{X+X'\given X'\},\cr
}$$
so $X+X'$ is independent of $X'$.

@@

\proof The triangle inequality gives $\dd\{X_1,X_1\} = 0$, and since $-x = x$ in $\FF_2^n$,
we also have $\dd\{X_1, - X_1\} = 0$. By the previous proposition, there is a coset $S$ of a
subgroup $H$ such that
$X_1$ has the same distribution as $U_S$. So
$$\dd\{X_1, U_H\} = \Eta\{U_S - U_H\} - {1\over 2}\Eta\{U_S\} - {1\over 2}\Eta\{U_H\} = 0,$$
since $U_S-U_H$ is uniform on $S$ and $|H| = |S|$. Then by the triangle inequality once again,
$\dd\{X_2, U_H\} = 0$ as well.\slug

@@

If we can prove this proposition, then we have Theorem~{\thmsubgroup} with a certain constant $C'$ which
shall be made explicit during the proof.

\medskip\noindent{\it Proof of Theorem~{\thmsubgroup} assuming Proposition~{\propmaintau}}.\enspace
We proved earlier that $\P(\FF_2^n)$ is a compact topological space, thus so is its Cartesian product with
itself. By Proposition~{\propruzsacontinuous}, the Ruzsa distance functional is continuous on this space,
which implies that $\tau$ is as well. Hence $\tau$ attains its infimum on the space $\P(\FF_2^n)^2$;
let $X_1$ and $X_2$ be two distributions such that $\tau\{X_1, X_2\}$ is minimal. By the contrapositive
of Proposition~{\propmaintau}, we must have $\dd\{X_1, X_2\} = 0$, so by Lemma~{\lemzerosubgroup},
there exists a subgroup $H$ of $\FF_2^n$ such that
$$\dd\{X_1, U_H\} = \dd\{X_2, U_H\} = 0.$$
This means that
$$\dd\{X_1^*, U_H\} \le \dd\{X_1^*, X_1\} + \dd\{ X_1, U_H\} = \dd\{X_1^*, X_1\}$$
and
$$\dd\{X_1^*, X_1\} \le \dd\{X_1^*, U_H\} + \dd\{U_H, X_1\} = \dd\{X_1^*, U_H\},$$
so $\dd\{X_1^*, U_H\} = \dd\{X_1^*, X_1\}$. The same holds for $X_2^*$ and $X_2$.
Then since $\dd\{X_1, X_2\} = 0$ and by minimality of $\tau\{X_1, X_2\}$, we obtain
$$\eqalign{
\eta\dd\{X_1^*, U_H\} + \eta\dd\{X_2^*, U_H\} &= \eta\dd\{X_1, U_H\} + \eta\dd\{X_2, U_H\} \cr
&= \tau\{X_1, X_2\} \cr
&\le \tau\{X_1^*, X_2^*\} \cr
&= \dd\{X_2^*, X_1^*\} + \eta\dd\{X_1^*, X_2^*\} + \eta\dd\{X_2^*, X_1^*\} \cr
&= (1+2\eta) \dd\{X_1^*, X_2^*\}. \cr
}$$
From our choice of $\eta = 1/9$ we can multiply both sides by $9$ to get
$$\dd\{X_1^*, U_H\} + \dd\{X_2^*, U_H\} \le 11\dd\{X_1^*, X_2^*\},$$
and furthermore, since
$$\bigl| \dd\{X_1^*, U_H\} - \dd\{X_2^*, U_H\} \bigr| \le \dd\{X_1^*, X_2^*\}$$
by two invocations of the triangle inequality, neither $\dd\{X_1^*, U_H\}$ nor
$\dd\{X_2^*, U_H\}$ can be greater than $6\dd\{X_1^*, X_2^*\}$.\slug

@@

\advsect Endgame

We now approach the phase of the proof which the authors of~\ref{pfr2023} theatrically call the `endgame.'
There is just one more lemma we need before we can give a full proof of Proposition~{\propmaintau}.

\edef\lemseventwo{\the\sectcount.\the\thmcount}
\proclaim Lemma~\advthm. Let $X_1$ and $Y_1$ be any $\FF_2^n$-valued random variables,
and let $T_1$, $T_2$, and $T_3$ be $\FF_2^n$-valued random variables
such that $T_1 + T_2 + T_3 = 0$ holds identically. Putting
$$\delta = \II\{T_1 : T_2\} + \II\{T_1:T_3\} + \II\{T_2 : T_3\}$$
and letting $\psi$ be the functional given by
$$\psi\{T_1', T_2'\}
= \dd\{T_1', T_2'\} + \eta\bigl( \dd\{X_1^*,T_1'\}-\dd\{X_1^*,X_1\}+\dd\{X_2^*,T_2'\}-\dd\{X_2^*, X_2\}\bigr),$$
there exist random variables $T_1'$ and $T_2'$ such that
$$\psi\{T_1', T_2'\} \le \biggl(1+{\eta\over 3}\biggr)\delta
+ {\eta\over 3}\sum_{i=1}^2\sum_{j=1}^3 \bigl(\dd\{X_i^*, T_j\}-\dd\{X_i^*, X_i\}\bigr)$$

\proof From the fact that any two of the $T_j$ determine the full triple $(T_1, T_2, T_3)$,
we have
$$\Eta\{T_1, T_2\} = \Eta\{T_1, T_3\} = \Eta\{T_2,T_3\} = \Eta\{T_1, T_2, T_3\}.$$
Then since $T_1 + T_2 = T_3$ we may apply Lemma~{\lematwopfr} with
$(X,Y,Z) = (T_1, T_2, T_3)$ to bound
$$\eqalign{
\sum_{t_3\in \FF_2^n} \pr\{T_3 &= t_3\} \dd\bigl\{ (T_1\given T_3 = t_3);
(T_2\given T_3 = t_3)\bigr\} \cr
&\le 2\II\{T_1:T_2\} + 2\Eta\{T_3\} - \Eta\{T_1,T_2\} \cr
&= 2\Eta\{T_1\} + 2\Eta\{T_2\} + 2\Eta\{T_3\} - 3\Eta\{T_1,T_2\} \cr
&= \II\{T_1,T_2\} + \II\{T_1, T_3\} + \II\{T_2,T_3\} \cr
&= \delta.
}$$
Letting $Z$ be a random variable independent of both $X_1^*$ and $X_2^*$, we apply Lemma~{\lemfiveone}
with $(X,Z,Y,W) = (X_1^*, Z, T_1, T_3)$ to obtain
$$\eqalign{
\sum_{t_3\in\FF_2^n}\pr\{T_3 = t_3\}\bigl(\dd\bigl\{X_1^*;(T_1\given T_3&=t_3)\bigr\}-\dd\{X_1^*, X_1\}\bigr) \cr
&= \dd\{ X_1^*, T_1\given T_3\} - \dd\{X_1^*, X_1\} \cr
&\le \dd\{X_1^*, T_1\} + {1\over 2}\II\{T_1:T_3\} - \dd\{X_1^*, X_1\}, \cr
}$$
and applying the same lemma with $(X,Z,Y,W) = (X_2^*, Z, T_2, T_3)$ yields
$$\eqalign{
\sum_{t_3\in\FF_2^n}\pr\{T_3 = t_3\}\bigl(\dd\bigl\{X_2^*;(T_2\given T_3&=t_3)\bigr\}-\dd\{X_2^*, X_2\}\bigr) \cr
&\le \dd\{X_2^*, T_2\} + {1\over 2}\II\{T_2:T_3\} - \dd\{X_2^*, X_2\}. \cr
}$$
Putting the three observations together, we have
$$\eqalign{
&\sum_{t_3\in\FF_2^n}\pr\{T_3 = t_3\}\psi\bigl\{(T_1\given T_3=t_3);(T_2\given T_3=t_3)\bigr\} \cr
&\hskip10pt = \sum_{t_3\in\FF_2^n}\pr\{T_3 = t_3\}\Bigl(\dd\bigl\{(T_1\given T_3=t_3);(T_2\given T_3=t_3)\bigr\} \cr
&\hskip103pt + \eta\dd\bigl\{X_1^*;(T_1\given T_3=t_3)\bigr\}+ \eta\dd\bigl\{X_2^*;(T_2\given T_3=t_3)\bigr\} \cr
&\hskip192pt - \eta\dd\{X_1^*, X_1\} - \eta\dd\{X_2^*, X_2\} \Bigr) \cr
&\hskip10pt \le \delta + \eta\bigl( \dd\{X_1^*, T_1\} - \dd\{X_1^*, X_1\}+\dd\{X_2^*, T_2\}-\dd\{X_2^*, X_2\}\cr
&\hskip180pt 
+ {1\over 2} \II\{T_1 : T_3\} + {1\over 2}\II\{T_2:T_3\} \bigr), \cr
}$$
Choosing some $t_3$ in the support of $T_3$ such that the value of $\phi$ inside
the sum is minimal and then setting $T_{1,3}' = (T_1\given T_3 = t_3)$ and
$T_{2,3}' = (T_2\given T_3 = t_3)$, we have
$$\eqalign{
\psi\bigl\{T_{1,3}', T_{2,3}'\bigr\} \le \delta + \eta\bigl( \dd\{X_1^*, T_1\} - \dd\{X_1^*, X_1\}
&+ \dd\{X_2^*, T_2\} - \dd\{X_2^*, X_2\} \cr
& + {1\over 2} \II\{T_1 : T_3\} + {1\over 2}\II\{T_2:T_3\} \bigr)\cr
}$$
We can now repeat this for all six permutations $(\alpha, \beta, \gamma)$ of $(1,2,3)$ and average the
corresponding bounds for $(T_\alpha, T_\beta, T_\gamma)$ to obtain
$$\eqalign{
{1\over 6} \sum_{(\alpha,\beta,\gamma)\in {\frak S}_3}
\psi\bigl\{T_{\alpha,\gamma}', T_{\beta,\gamma}'\bigr\}
\le \delta &- \eta\dd\{X_1^*, X_1\} - \eta\dd\{X_2^*, X_2\} \cr
&+ {\eta\over 6}\sum_{(\alpha,\beta,\gamma)\in {\frak S}_3}
\bigl( \dd\{X_1^*, T_\alpha\} + \dd\{X_2^*, T_\beta\} \cr
&\hskip68pt+ {1\over 2} \II\{T_\alpha : T_\gamma\} + {1\over 2}\II\{T_\beta:T_\gamma\} \bigr).\cr
}$$
Since each of $\{1,2,3\}$ appears twice as $\alpha$ and twice as $\beta$, we have
$$\sum_{(\alpha,\beta,\gamma)\in {\frak S}_3} \bigl(\dd\{X_1^*, T_\alpha\} + \dd\{X_2^*, T_\beta\}\bigr)
= 2\sum_{i=1}^2\sum_{j=1}^3 \dd\{X_i^*, T_j\}$$
and
$$\sum_{(\alpha,\beta,\gamma)\in {\frak S}_3}
\biggl({1\over 2} \II\{T_\alpha : T_\gamma\} + {1\over 2}\II\{T_\beta:T_\gamma\}\biggr) = 2\delta.$$
Consequently, the average of
$\psi\bigl\{T_{\alpha,\gamma}', T_{\beta,\gamma}'\bigr\}$
over all permutations $(\alpha,\beta,\gamma)$ is bounded above by
$$\biggl(1+{\eta\over 3}\biggr)\delta
- \eta\dd\{X_1^*, X_1\} -\eta\dd\{X_2^*, X_2\} + {\eta\over 3}\sum_{i=1}^2\sum_{j=1}^3 \dd\{X_i^*, T_j\},$$
so the result follows by letting $(T_1', T_2')$ equal the pair $(T_{\alpha,\gamma}', T_{\beta,\gamma}')$
for the choice of $(\alpha,\beta,\gamma)$ that minimises
$\psi\bigl\{T_{\alpha,\gamma}', T_{\beta,\gamma}'\bigr\}$.\slug

We are now able to prove Proposition~{\propmaintau}, which we shall restate in the contrapositive.
Just as in previous sections, the functional $\tau$ depends on the fixed random variables $X_1^*$ and $X_2^*$
as well as the choice of constant $\eta = 1/9$.

\proclaim Proposition~{\propmaintau}$\mathbold'$. Let $X_1$ and $X_2$ be $\FF_2^n$-valued random variables
with the property that $\tau\{X_1', X_2'\} \ge \tau\{X_1, X_2\}$ for all random variables
$X_1'$ and $X_2'$ on $\FF_2^n$. Then $\dd\{X_1, X_2\} = 0$.
